Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2012-01-06 ltx/paper.tex
added more to relevant work

\documentclass[smallcondensed, draft]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]

\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
\titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}

\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
\authorrunning{Willie Neiswanger and Frank Wood}

\institute{$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
Tel.: 503-464-6152.
\email{wdn2101@columbia.edu}
\and
$^{\dagger}$Columbia University, Department of Statistics. 
Tel.: 212-851-2150. 
\email{fwood@stat.columbia.edu}}

\date{}

@@

@@

@@

@@

Localization is related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation and pixel clustering \willie{must cite here}. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

@@

In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times \willie{cite paper(s) for each of these}.

In the past five years, there has been a push for methods that incorporate detection and tracking of multiple objects to allow for full systems capable of autonomously handling real-world video datasets.
\willie{Describe tracking by detection, more on blob tracking, the other few methods i've found for automated joint detection and tracking}

@@

% %  ------------------------------------------------------------------------------------
% % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).

@@

% % Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking). Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking. Work of Pece, MTT of multiple targets work(?). The work presented in this paper most closely resembes the topics listed here. Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.
% %  ------------------------------------------------------------------------------------

@@

@@

@@

We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). %An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}...The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.

@@

@@

@@

@@

@@

@@

% An alternative formultion of the DPM is known as the Chinese resturant process (CRP) sampling representation. 
If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$. Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned. We can now formulate the Chinese restaurant process, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$). The Chinese restaurant process allows samples to be drawn from the conditional distribution over the assignment variables $c_{i}$, and can be expressed as
\begin{align}
\begin{split}
\label{crp_rep}
	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
	P(c_{i} \neq c_{j} \text{  for all  } j<i) &= \frac{\alpha}{i-1+\alpha}
\end{split}
\end{align}
where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$. A modified Chinese restaurant process will be used to draw samples from the conditional distribution over the assignment variables when performing inference on the GPUDDPM (Section~\ref{sec:gpuddpm}).

@@

@@

Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (see \eqref{crp_rep}). Hence, to account for varying numbers of clusters---and in particular, to allow clusters to diminish in size (i.e. reduce the number of assigned observations) and even die---there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered ``removed'' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.

The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion. Specifically, each (remaining) assignment is removed from its cluster with probability $\rho$. The size $m_{k,t}$ of cluster $k$ at time $t$ can be computed by removing elements from cluster $k$ at the previous time and thus diminishing the cluster's size at time $t-1$. It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the cluster size $m_{k,t}$ by $r$. % This can be written as
% \begin{equation}
% \label{del_step}
% m_{k,t} | m_{k,t-1}   \sim   m_{k,t-1}-\text{Binomial}(m_{k,t-1}, \rho)
% \end{equation}
For each time step, after all observations are given an assignment, this deletion procedure is carried out. Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$, is dependent upon the cluster size at the previous time step and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$. We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL, which we define to be
\label{del_step}
\text{DEL} & (\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) :=
\left( m_{k,t-1} - \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) \right) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
$\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise. This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}. The Chinese restaurant process is used to sample assignment variables during inference procedures (described in Section~\ref{sec:inference}). We define a distribution used in these procedures, CRP, to be

@@

$\forall i \in \{1, \ldots, N \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$. Note the similarity between the above definition and the Chinese resturant process of the DPM given by \eqref{crp_rep}. The deletion and Chinese resturant procedures together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. Using the DEL and CRP distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,  % and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,

@@

@@

@@

In the following sections, a particular extraction method for generating data from videos and a specific formulation of the model described in Section~\ref{sec:modeloverview} are given. Inference schemes on this model, which when carried out allow for unsupervised detection and tracking of multiple arbitrary objects, are also described.

@@

@@

@@

In the following experiments, the hue component of a pixel's h-s-v value is recorded from each pixel surrounding a given $\bold{x}^{s}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$). Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$. Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated (\willie{cite mckenna, other papers here}).

@@

@@

@@

@@

The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. Object apperance parameters include\\\\
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
$\bold{q}_{0}$  &  scale factor of the prior on the multinomial counts
\end{tabular}
\end{center} \vspace{5mm}
The following parameter dictates characteristics involving the movement of objects
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
\end{tabular}
\end{center} \vspace{5mm}
Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
\end{tabular}
\end{center}
\willie{Discuss choosing parameters}

@@

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}. This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions. To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time. Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster . At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. In particular, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as
\begin{equation}
\label{compute_clust_size}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
\begin{equation}
\label{del_rho_form}
l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
\end{equation}
The MCMC sampler described in this section operates in this alternate representation, which we can write, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$, as
% \begin{align}
% \begin{split}
% 	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
% 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
% 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
% 	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
% \begin{cases}
% 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
% 	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
% \end{cases}
% \end{split}
% \end{align}
% where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
\begin{align}
\begin{split}
d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
\theta_{k, t} | \theta_{k, t-1}  &\sim
\begin{cases}
P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
\mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
\end{cases} \\
c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  F(\theta_{c_{i,t}, t})
\end{split}
\end{align}
where $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.

>>>>

% header stuff
% ------------

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

% title and author related
% ------------------------

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}

\titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}

% \subtitle{}

\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}

\authorrunning{Willie Neiswanger and Frank Wood}

\institute{ $^{*}$Columbia University, Department of Applied Math and Applied Physics. Tel.: 503-464-6152. \email{wdn2101@columbia.edu} \and $^{\dagger}$Columbia University, Department of Statistics. Tel.: 212-851-2150. \email{fwood@stat.columbia.edu}}

\date{}  % add submit date when submitted

@@

% Abstract
% --------

@@

% introduction
% ------------

@@

% related work
% ------------

@@

Localization is related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving model-based clustering of individual pixels, local image texture representations, or other patterns found in image regions \cite{jain1997object, fei2005bayesian, sivic2005discovering}. These works usually aim to automatically segment images into meaningful regions. Note that the popular phrase `object detection' is sometimes used to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection, which incorporates both the task of determining what constitutes distinct objects and localizing these objects), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

@@

In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times. Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object. To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers. An approach to this problem involving the use of mixture models has also been studied by \cite{vermaak_2003}; this paper found success in maintaining isolation of distinct objects through the use of a nonparametric mixture model.

In the past five years, there has been a push for methods that attempt to provide general algorithms for the detection and tracking of multiple objects in videos. Blob tracking, a basic method attempting to carry out this goal, has been in use for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible. Blob tracking-based methods run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, there exists object occlusion, 
\willie{Describe tracking by detection, more on blob tracking, the other few methods i've found for automated joint detection and tracking}

% extra background that I decided to cut (for now)
% ------------------------------------------------

@@

% more background that I decided to cut (for now)
% -----------------------------------------------

% % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).

@@

% % Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking). Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking. Work of Pece, MTT of multiple targets work(?). The work presented in this paper most closely resembes the topics listed here. Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.

% Contributions (decided to cut for now)
% --------------------------------------

@@

% general model overview section
% ------------------------------

@@

% general data extraction section
% -------------------------------

@@

We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).

% this is the pretty worms figure that we decided to cut
% ------------------------------------------------------

%An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}...The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.

@@

% section providing background to introduce DPMs
% ----------------------------------------------

@@

% defining the finite mixture model
% ---------------------------------

@@

% defining the Bayesian mixture model
% -----------------------------------

@@

% defining Dirichlet processes
% ----------------------------

@@

% defining the DPM
% ----------------

@@

If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$. Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned. We can now formulate the Chinese restaurant process, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$). The Chinese restaurant process allows samples to be drawn from the conditional distribution over the assignment variables $c_{i}$, and can be expressed as
\begin{align}
\begin{split}
\label{crp_rep}
	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
	P(c_{i} \neq c_{j} \text{  for all  } j<i) &= \frac{\alpha}{i-1+\alpha}
\end{split}
\end{align}
where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$. A modified Chinese restaurant process will be used to draw samples from the conditional distribution over the assignment variables when performing inference on the GPUDDPM (Section~\ref{sec:gpuddpm}).

% section providing background to introduce DDPMs
% -----------------------------------------------

@@

% defining the GPUDDPM
% --------------------

@@

Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (as per \eqref{crp_rep}). Hence, to account for varying numbers of clusters---and in particular, to allow clusters to diminish in size (i.e. to diminish in number of assigned observations) and even ``die off"---there is a deletion procedure that allows observations to be considered removed from their assigned clusters at a given time step. This procedure is introduced and described in detail in \cite{caron_2007}.

The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion. Specifically, each remaining assignment is removed from its cluster with probability $\rho$.  It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the previous-time cluster size $m_{k,t-1}$ by $r$. At each time step, this deletion procedure is carried out on all existing assignments, including those at the current time. Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$ is dependent upon the cluster size at the previous time step, $m_{k,t-1}$, and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$. We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL, which we define to be
\begin{equation}
\begin{split}
\label{del_step}
\text{DEL} & (\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) :=
\left( m_{k,t-1} - \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) \right) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
\end{split}
\end{equation}
$\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, $K_{t}$ is the number of clusters at time $t$, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise. This process adheres to what \cite{caron_2007} calls a ``uniform deletion strategy'' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}.

The Chinese restaurant process, defined in \eqref{crp_rep}, is used to sample assignment variables during inference procedures (described in Section~\ref{sec:inference}). Based on this definition, we define the distribution CRP to be

@@

$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$. Note the similarity between the above definition and the Chinese resturant process of the DPM given by \eqref{crp_rep}. The deletion and Chinese resturant procedures together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. Using the DEL and CRP distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,  % and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,

@@

% section making explicit connection between model and video tracking
% -------------------------------------------------------------------

@@

% section providing specific model
% --------------------------------

@@

In the following sections, a particular extraction method for generating data from videos and a specific formulation of the general model described in Section~\ref{sec:modeloverview} are given. Inference schemes on this model, which when carried out allow for unsupervised detection and tracking of multiple arbitrary objects, are also described.

% gives specific data extraction method (frame differencing)
% ----------------------------------------------------------

@@

% introduce specific model for experiments
% ----------------------------------------

@@

% data in experiments
% -------------------

@@

In the following experiments, the hue component of a pixel's hue-saturation-value (h-s-v) tuple is recorded from each pixel surrounding a given $\bold{x}^{s}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$). Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$. Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated (\willie{cite mckenna, other papers here}).

% likelihood / object appearance model
% ------------------------------------

@@

% base distribution / appearance prior
% ------------------------------------

@@

% transition kernel
% -----------------

@@

% recap of parameters
% -------------------

@@

The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. \\ \\
Object appearance parameters include:
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
$\bold{q}_{0}$  &  scale factor of the prior on the multinomial counts
\end{tabular}
\end{center} \vspace{3mm}
The following parameter dictates characteristics involving the movement of objects
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
\end{tabular}
\end{center} \vspace{3mm}
Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
\begin{center}
\begin{tabular}[c]{l p{10cm}}
$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
\end{tabular}
\end{center}
\willie{Discuss choosing parameters}

% inference section
% -----------------

@@

% MCMC inference section
% ----------------------

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}. This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions. To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time. Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster . At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. In particular, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as
\label{compute_clust_size}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
\label{del_rho_form}
l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
The MCMC sampler described in this section operates in this alternate representation, which we can write, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$, as
% 	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
% 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
% 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
% 	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
% 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
% 	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
% \end{cases}
% where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
\mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
% c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  F(\theta_{c_{i,t}, t})
where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}. This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
        \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
        \label{fig:gpuddpm_gm_2}
